\begin{abstract}
\noindent{\bf\zihao{-4}摘要:}尾矿库是矿山企业最大的危险源之一，加强尾矿库的安全检测是保障矿山企业正常生产的重要手段和方法。
浸润线的高低反映了尾矿库的安全状态，是尾矿库安全监测的重要内容。
传统的浸润线检测手段主要包括测压管、传感器等。这些方法效率低、耗费高，严重限制了尾矿库安全监测水平的提高。
地质雷达作为一种无损检测手段，具有高效、低耗费、检测灵活的特点，在尾矿库浸润线检测方面有广阔的应用前景。
本文首先结合水利大坝和地下水埋深检测经验，分析得出尾矿库探测介质主要成分为尾矿砂，可以当做砂土处理；
然后，本文分析得出尾矿库探测的主要影响因子为含水量，借助包气带水分布模型和Toop公式研究了尾矿浸润线过渡区对探测精度的影响，认为当过渡区厚度小于波长时探测是可行的；
其次，本文根据尾矿库的介电模型，利用GPRSIM和MATGPR对浸润线探测做了正反演分析，得出浸润线的图像特征为同相轴组成的连续反射带；
最后，本文对尾矿库进行了实地探测，分析其图像发现多层状连续减弱的反射带，通过其振幅衰减可以识别出浸润线位置，研究认为地质雷达探测浸润结果是可靠的。


\keywords{浸润线;探测;地质雷达;正演}
\end{abstract}
\begin{englishabstract}
\noindent{\bf\zihao{-4}Abstract：}Tailings is the biggest hazards for mining enterprises, strengthen the safety monitoring is to ensure the tailings mining enterprise normal production of premise. Infiltrating line reflects the tailings safe state, is one important part of the safety monitoring for tailings. The traditional infiltration detection methods are mainly including piezometer tube, sensors, etc. These methods the efficiency is low, the cost is high, severely restricting tailings to raise the level of safety monitoring. GPR as a nondestructive testing method, with high efficiency and low cost, testing flexible characteristics, in the inspection of the tailings infiltration line have broad application prospects. This article tries to water conservancy dam and groundwater depth detection experience, use of geological radar detection tailings infiltrating line. First of all, article with water bag with gas distribution model and the Toop formula analyzes tailings detection conditions, fitting the dielectric model curve, analyzes the feasibility of detecting; Second, according to the tailings dielectric model, we use GPRSIM and MATGPR for forward modeling analysis, studied the features of infiltration line images; Finally, this paper analyzes the tailings infiltrating line detection results In engineering application, proving that the GPR is effective way for Infiltrating line detection

\englishkeywords{seepage line, detection, Ground Penetrating Radar, forward modeling}
\end{englishabstract} 